```{r include=FALSE}
page_title <- "Coho FRAM Applications"
```

---
title: `r page_title`
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE, results = 'hide', warning = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.height=9, fig.width=9)
```

# 1. Pre-season model input development

Each year fishery managers convene to plan the next season of salmon fisheries. This pre-season planning process is generally known as the "North of Falcon" (<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#North_of_Falcon_(NOF)" target = "_blank" title = "NOF: North of Falcon">NOF</a>) process, referencing the region north of Cape Falcon, Oregon which marks the southern border of active management for Washington salmon stocks. The <span title = "NOF: North of Falcon">NOF</span> planning process coincides with the March and April meetings of the Pacific Fishery Management Council (<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pacific_Fishery_Management_Council_(PFMC)" target = "_blank" title = "PFMC: Pacific Fishery Management Council">PFMC</a>), but includes additional public meetings as well as government to government negotiations. Final <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Fishery_Regulation_Assessment_Model_(FRAM)" target = "_blank" title = "FRAM: Fishery Regulation Assessment Model">FRAM</a> model runs are a joint exercise which encompass fisheries from both the <span title = "PFMC: Pacific Fishery Management Council">PFMC</span> and <span title = "NOF: North of Falcon">NOF</span> pre-season negotiations and are usually completed at the end of the April <span title = "PFMC: Pacific Fishery Management Council">PFMC</span> meeting.

A wide range of model inputs are compiled and processed for <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> pre-season model runs. Model inputs are received from the respective government agencies/tribes in Canada (Department of Fisheries and Oceans, <span title = "DFO: Department of Fisheries and Oceans">DFO</span>), Washington state (Washington Department of Fish and Wildlife, <span title = "Washington Department of Fish and Wildlife">WDFW</span>), twenty Western Washington treaty Indian tribes (member tribes of the Northwest Indian Fisheries Commission, <span title = "NWIFC: Northwest Indian Fisheries Commission">NWIFC</span>), the Columbia River Intertribal Fish Commission (<span title = "CRITFC: Columbia River Intertribal Fish Commission">CRITFC</span>), Oregon (Oregon Department of Fish and Wildlife, <span title = "ODFW: Oregon Department of Fish and Wildlife">ODFW</span>), and California (California Department of Fish and Game, <span title = "CDFG: California Department of Fish and Game">CDFG</span>).

The process of developing <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> model inputs begins with the forecasting of <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Hatchery" target = "_blank" title = "Hatchery">hatchery</a> and wild/<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Natural" target = "_blank" title = "Natural">natural</a> <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Stock" target = "_blank" title = "Stock">stock</a> abundances during January-February.  From February through April, fishery inputs for <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> are generally developed and updated through the <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pre-season" target = "_blank" title = "Pre-season">pre-season</a> process following technical staff discussions and manager agreement for ocean fisheries through the <span title = "PFMC: Pacific Fishery Management Council">PFMC</span> and for Puget Sound tribal and state fisheries through <span title = "NOF: North of Falcon">NOF</span>.  Fishery-related mortality <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Parameter" target = "_blank" title = "Parameter">parameters</a>, such as release mortality rates, drop-off, and <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Mark-selective_Fishery_(MSF)" target = "_blank" title = "MSF: Mark-selective fishery">mark-selective fishery</a> parameters, are reviewed and confirmed at the start of the annual management cycle.   

## 1.1. Stock Abundance

A variety of methods are used to forecast Coho abundances. These forecasts are usually developed by local/regional technical staff during one or more technical meetings where relevant forecasting information is exchanged.  Final agreed-to forecasts are published in the <span title = "PFMC: Pacific Fishery Management Council">PFMC</span> Salmon Technical Team <a href = "https://www.pcouncil.org/documents/2021/03/2021-preseason-report-i.pdf/" target = "_blank" title = "Pre-season Report 1">Pre-season Report 1</a>; <a href = "https://www.dfw.state.or.us/agency/commission/minutes/19/03_Mar/E/Attachment%204_Preseason%20Report%20I.pdf" target = "_blank" title = "Stock Abundance Analysis and Environmental Assessment Part 1">Stock Abundance Analysis and Environmental Assessment Part 1</a> (Stock Assessment and Fishery Evaluation; <span title = "SAFE: Stock Assessment and Fishery Evaluation">SAFE</span>). Columbia River, Oregon, and California stock forecasts are provided through the Oregon Production Index Technical Team (<span title = "OPITT: Oregon Production Index Technical Team">OPITT</span>) to <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> modelers. Although Alaskan stocks and fisheries are in the Coho <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> model, they are not modeled dynamically pre-season and instead use static scalers to represent a proxy of stock abundances and base period fishing. Canadian stock abundances are provided by <span title = "Pacific Salmon Commission">PSC</span> Coho Technical Team and Southern Panel representatives through an annual exchange meeting between the U.S. and Canada. This exchange meeting regularly occurs just prior to the first <span title = "NOF: North of Falcon">NOF</span> salmon co-manager meeting in mid-March (between March and April <span title = "PFMC: Pacific Fishery Management Council">PFMC</span> meetings). Until these Canadian stock inputs are available, values from the previous year must be used which creates greater uncertainty during the annual planning process.

<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Abundance_Forecast" target = "_blank" title = "Abundance forecasts">Abundance forecasts</a> vary in units of measure.  For example, there are forecasts of salmon returning to a terminal area (which implies some accounting for <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pre-terminal_Fishery" target = "_blank" title = "Pre-terminal fishery">pre-terminal fishery</a> levels), forecasts of ocean abundance (which is commonly fishery impacts plus <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Escapement" target = "_blank" title = "Escapement">escapement</a>), and forecasts of abundances prior to any fishing impacts (which includes <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Natural_Mortality" target = "_blank" title = "Natural mortality">natural mortality</a> and <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Non-landedNon-retention_Mortality" target = "_blank" title = "Non-landed mortality">non-landed</a> fishery related mortality). Each of these different types of forecasts need to be converted to the “unit of measure” used by <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span>, which is the abundance of each stock prior to fishing vulnerability and natural mortality.  This starting Coho <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> unit of measure is characterized as January-age-3 fish (<span title = "JA3: January age 3 abundance">JA3</span>). <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> stock abundances are input as a stock recruit scaler, which is calculated as the forecasted number of fish divided by the <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Base_Period" target = "_blank" title = "Base period">base period</a> average abundance for each stock.  
 
Many U.S. domestic and Pacific Salmon Treaty (<span title = "PST: Pacific Salmon Treaty">PST</span>) International Coho management objectives are specified in ocean age-3 abundance units (<span title = "OA3: Ocean age-3 abundance">OA3</span>, fishery impacts plus escapement).  Thus, most of the forecasts are produced in terms of ocean age-3 abundance, which are then expanded by 1.2317 to account for natural mortality and provide an estimate of age-3 fish abundance in January of the fishing year (<span title = "OA3: Ocean age-3 abundance">OA3</span> * 1.2317 = <span title = "JA3: January age-3 abundance">JA3</span>).  The value 1.2317 is a result of a two-step conversion due to a prior December age-2 (<span title = "DA2: December age-2 abundance">DA2</span>) unit of measure used in modeling before <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> (1.2317 = 1.333 (convert <span title = "OA3: Ocean age-3 abundance">OA3</span> to <span title = "DA2: December age-2 abundance">DA2</span>) * 0.924 (convert <span title = "DA2: December age 2 abundance">DA2</span> to <span title = "JA3: January age 3 abundance">JA3</span>)).  Any non-landed fishery related mortality that occurs is ignored in this ocean abundance-to-total abundance <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> conversion step.  

## 1.2. Fisheries

For Puget Sound and the Washington coast, data and methods used to generate fishery inputs for <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> and <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Terminal_Area_Management_Module_(TAMM)" target = "_blank" title = "TAMM: Terminal Area Management Module">TAMM</a> are shared, discussed, and agreed-to by salmon <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Co-managers" target = "_blank" title = "Co-managers">co-manager</a> technical staff in the <span title = "NOF: North of Falcon">NOF</span> annual process. Fishery inputs are compiled by the <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> modelers using submissions from the respective tribes (ceremonial & subsistence, commercial and test fisheries) and <span title = "Washington Department of Fish and Wildlife">WDFW</span> (recreational, commercial and test fisheries). In general, fishery inputs can be either provided as scalers, quotas, or <span title = "TAMM: Terminal Area Management Module">TAMM</span> harvest rates (mark-selective or non-selective). For the calculation of recreational Puget Sound marine fishery inputs, an external file is used. Based on historical catch record card and creel survey-based estimates of mortalities, recreational fishery inputs are calculated depending on the fishery type (mark-selective, non-selective, catch and release and non-retention), months or days open for the respective fishery, and any regional or fishery specific adjustments (e.g. two pole adjustment). Ocean fishery inputs for <span title = "PFMC: Pacific Fishery Management Council">PFMC</span>-managed fisheries are provided to the <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> modelers by staff from Washington coastal Indian tribes, <span title = "CRITFC: Columbia River Intertribal Fish Commission">CRITFC</span>, <span title = "Washington Department of Fish and Wildlife">WDFW</span>, <span title = "ODFW: Oregon Department of Fish and Wildlife">ODFW</span> and <span title = "CDFG: California Department of Fish and Game">CDFG</span>.  

<span title = "PFMC: Pacific Fishery Management Council">PFMC</span>-managed Coho retention fisheries are modeled as landed fish quotas or mark-selective fishery quotas (i.e., retained <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Marked" target = "_blank" title = "marked">marked</a> fish).  <span title = "PFMC: Pacific Fishery Management Council">PFMC</span>-managed Coho non-retention fisheries are modeled using external estimates of mortalities generated from historical Coho to Chinook ratios of landings when retention of both species was allowed.  In some fisheries, like the <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Troll_Fishery" target = "_blank" title = "Troll fishery">troll fisheries</a> South of Cape Falcon, these external mortality estimates are adjusted downward to account for shifts in effort away from the species that cannot be retained.  

Fisheries outside of <span title = "PFMC: Pacific Fishery Management Council">PFMC</span> jurisdiction are modeled using quotas as a landed catch expectation, as catch (or occasionally effort) scalers, or as terminal area <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Harvest_Rate_(HR)" target = "_blank" title = "HR: Harvest rate">harvest rates</a> used during <span title = "TAMM: Terminal Area Management Module">TAMM</span> processing.  Canadian fisheries are modeled using <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Fishery_Scaler" target = "_blank" title = "Fishery scaler">fishery scalers</a> generated from <a href = https://framverse.github.io/fram_doc/calcs_glossary.html#Post-season" target = "_blank" title = "Post-season">post-season</a> <span title = "PST: Pacific Salmon Treaty">PST</span> Coho Technical Committee (<span title = "CoTC: Coho Technical Committee">CoTC</span>) Coho <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> modeling exercises, which are intended to represent similar fishery expectations or regulations in the upcoming fishing year.  Alaskan fisheries are modeled pre-season using static scalers to represent proxy stock abundances relative to Coho <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> base period fishing. 

# 2. Post-season Model Usage

Although Coho <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> is primarily used for pre-season fishery impact assessment, it can also be used in a post-season context.  These post-season model runs are primarily conducted for two purposes: as a tool to evaluate the model's performance by comparing model estimates to independently derived estimates, and to evaluate the performance of the fishery management system towards meeting conservation objectives for key stocks. Post-season model runs are occasionally used as a "one-stop-shop" to get a historic time series of Coho escapements or pre-fishing abundances of hatchery and natural stock units for the purpose of conducting population studies or evaluating forecasts.

This context uses the <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> model to estimate what occurred in a previous year based on observed input values and the associated estimated abundances. Post-season <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> model runs use the “backwards” <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Backwards_FRAM_(BkFRAM)" target = "_blank" title = "BkFRAM: Backwards FRAM">BkFRAM</a> utility that finds <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> starting abundances using inputs of observed estimates of escapements and fishery catches (landed and non-retention). Post-season model runs are conducted by the Pacific Salmon Commission (<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pacific_Salmon_Commission_(PSC)" target = "_blank" title = "PSC: Pacific Salmon Commission">PSC</a>) CoTC every year during the <span title = "PSC: Pacific Salmon Commission">PSC</span> annual meeting (February) and used to evaluate management obligations under the <span title = "PST: Pacific Salmon Treaty">PST</span>. Post-season model runs are often lagged by several years due to data availability issues (e.g., in 2021 for fishery year 2019).  

<br>

**To cite this page:** <br>
`r paste("Salmon modeling and analysis workgroup.", format(Sys.Date(), "%Y."), page_title, "*in* FRAM Documentation. https://framverse.github.io/fram_doc/ ", paste("built", format(Sys.Date(), "%B %d, %Y.")))`
