```{r include=FALSE}
page_title <- "Chinook FRAM Applications"
```

---
title: `r page_title`
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE, results = 'hide', warning = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.height=9, fig.width=9)
```

# 1. Pre-season Model Input Development

Each year fishery managers convene to plan the next fishing season of salmon fisheries. This pre-season planning process is generally known as the "North of Falcon" (<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#North_of_Falcon_(NOF)" target = "_blank" title = "NOF: North of Falcon">NOF</a>) process, referencing the region north of Cape Falcon Oregon which marks the southern border of active management for Washington salmon stocks. The <span title = "NOF: North of Falcon">NOF</span> planning process coincides with the March and April meetings of the Pacific Fishery Management Council (<span title = "PFMC: Pacific Fishery Management Council">PMFC</span>) but includes additional public meetings as well as government to government negotiations. Final <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Fishery_Regulation_Assessment_Model_(FRAM)" target = "_blank" title = "FRAM: Fishery Regulation Assessment Model">FRAM</a> model runs are a joint exercise which encompass fisheries from both the <span title = "PFMC: Pacific Fishery Management Council">PMFC</span> and <span title = "NOF: North of Falcon">NOF</span> pre-season negotiations and are usually completed at the end of the April <span title = "PFMC: Pacific Fishery Management Council">PMFC</span> meeting.

A wide range of model inputs are compiled and processed for <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> pre-season model runs. Model inputs are received from the respective government agencies/tribes in Canada (Department of Fisheries and Oceans, <span title = "DFO: Department of Fisheries and Oceans">DFO</span>), Alaska (Alaska Department of Fish and Game, <span title = "ADFG: Alaska Department of Fish and Game">ADFG</span>), Washington state (Washington Department of Fish and Wildlife, <span title ="WDFW: Washington Department of Fish and Wildlife">WDFW</span>), twenty Western Washington treaty Indian tribes (member tribes of the Northwest Indian Fisheries Commission, <span title = "NWIFC: Northwest Indian Fisheries Commission">NWIFC</span>), the Columbia River Intertribal Fish Commission (<span title = "CRITFC: Columbia River Intertribal Fish Commission">CRITFC</span>), Oregon (Oregon Department of Fish and Wildlife, <span title = "ODFW: Oregon Department of Fish and Wildlife">ODFW</span>), and California (California Department of Fish and Game, <span title = "CDFG: California Department of Fish and Game">CDFG</span>). 

The process of developing the <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> model inputs begins with the forecasting of <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Hatchery" target = "_blank" title = "Hatchery">hatchery</a> and wild/<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Natural" target = "_blank" title = "Natural">natural</a> <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Stock" target = "_blank" title = "Stock">stock</a> abundances during January-February.  From February through April, fishery inputs for <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> are generally developed and updated through the <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pre-season" target = "_blank" title = "Pre-season">pre-season</a> process following technical staff discussions and manager agreement for ocean fisheries through the <span title = "PFMC: Pacific Fishery Management Council">PMFC</span> and for 
Puget Sound tribal or state fisheries through <span title = "NOF: North of Falcon">NOF</span>.  Fishery-related mortality <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Parameter" target = "_blank" title = "Parameter">parameters</a>, such as release mortality rates, drop-off, sublegal/legal ratios, and <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Mark-selective_Fishery_(MSF)" target = "_blank" title = "MSF: Mark-selective fishery">mark-selective fishery</a> parameters, are reviewed and confirmed at the start of the annual management cycle. 

## 1.1. Stock Abundance

A variety of methods are used to forecast Chinook abundances.  These forecasts are usually developed by local/regional technical staff during one or more technical meetings where relevant forecasting information is exchanged.  Final agreed-to forecasts are published in the <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pacific_Fishery_Management_Council_(PFMC)" target = "_blank" title = "PFMC: Pacific Fishery Management Council">PFMC</a> Salmon Technical Team <a href = "https://www.pcouncil.org/documents/2021/03/2021-preseason-report-i.pdf/" target = "_blank" title = "Pre-season Report 1">Pre-season Report 1</a>; <a href = "https://www.dfw.state.or.us/agency/commission/minutes/19/03_Mar/E/Attachment%204_Preseason%20Report%20I.pdf" target = "_blank" title = "Stock Abundance Analysis and Environmental Assessment Part 1">Stock Abundance Analysis and Environmental Assessment Part 1</a> (Stock Assessment and Fishery Evaluation; <span title = "SAFE: Stock Assessment and Fishery Evaluation">SAFE</span>).

<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Abundance_Forecast" target = "_blank" title = "Abundance forecasts">Abundance forecasts</a> vary in units of measure.  For example, there are forecasts of salmon returning to a terminal area (which implies some accounting for <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pre-terminal_Fishery" target = "_blank" title = "Pre-terminal fishery">pre-terminal fishery</a> levels), forecasts of ocean abundance (which is commonly fishery impacts plus <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Escapement" target = "_blank" title = "Escapement">escapement</a>), and forecasts of abundances prior to any fishing impacts (which includes <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Natural_Mortality" target = "_blank" title = "Natural mortality">natural mortality</a> and <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Non-landedNon-retention_Mortality" target = "_blank" title = "Non-landed mortality">non-landed</a> fishery related mortality).  In the case of Chinook, the forecasts of fish returning to the terminal area need to account for pre-terminal fishing impacts or impacts that occurred in previous seasons.

Although abundance forecasts vary in units of measure, the preferred units for Chinook are either Terminal Run Size (<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Terminal_Run_Size" target = "_blank" title = "TRS: Terminal run size">TRS</a>) or Extreme Terminal Run Size (<span title = "ETRS: Extreme Terminal Run Size">ETRS</span>). The <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> stock terminal run sizes are imported into <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> via an external ChinRSScalarsMR.xlsm file and the <span title = "TAMM: Terminal Area Management Module">TAMM</span> stock terminal run sizes are copied from the ChinRSScalarsMR.xlsm file into the "Input Page" sheet of the <span title = "TAMM: Terminal Area Management Module">TAMM</span> file. TRS forecasts are then run through a "backwards" model run using the <a title = "BkFRAM: Backwards FRAM" href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Backwards_FRAM_(BkFRAM)" target = "_blank">BkFRAM</a> utility (<a href = "https://framverse.github.io/fram_doc/calcs_data_chin.html#6_Backwards_FRAM" target = "_blank" title = "Chinook Model Detail Chapter 6: BkFRAM">Chinook Model Detail Chapter 6</a>), which then generates the starting cohorts for the abundance at age of each stock prior to fishing and natural mortality. This model run uses agreed-to average observed fishery input data from a set of prior years to represent pre-terminal fishery levels when deriving the starting <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> abundances from the original terminal run size forecasts. This particular model run is commonly referred to as the initial "average run" prior to any regular pre-season modeling.  Abundance of age 2 fish are either forecasted separately and provided as a model input or can be generated using an “age 2 from 3” utility feature in <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> (see <a href = "https://framverse.github.io/fram_doc/user_prereq.html" target = "_blank">FRAM User Manual</a>).

## 1.2. Fisheries

Chinook <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> is the primary tool used to plan and assess pre-season fishery packages for ocean fisheries through the <span title = "PFMC: Pacific Fishery Management Council">PFMC</span> process and for Puget Sound fisheries through the <span title = "NOF: North of Falcon">NOF</span> process.  While a variety of methods are used to model fisheries in a pre-season context, the methods used for each fishery are generally consistent between years.

Puget Sound fishery inputs for <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> and <span title = "TAMM: Terminal Area Management Module">TAMM</span> are shared, discussed, and agreed-to by salmon co-manager technical staff in the <span title = "NOF: North of Falcon">NOF</span> annual process. Puget Sound fisheries are modeled using a variety of the available <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> methods. <span title = "PFMC: Pacific Fishery Management Council">PMFC</span> managed ocean fisheries are modeled by the Salmon Technical Team (<span title = "SST: Salmon Technical Team">STT</span>). For <span title = "PFMC: Pacific Fishery Management Council">PMFC</span> fisheries north of Cape Falcon, which are managed based on a total allowable catch limit, quota inputs are used for both commercial troll and recreational fisheries.

Fisheries south of Cape Falcon are modeled using fishery scalers, as these fisheries are generally managed as a season. The fishery scalers are calculated as the expected number of vessel fishing days for troll fisheries or angler trips for recreational fisheries divided by the base period average effort levels. Effort expectations for each pre-season fishery package are provided from output of the Klamath Ocean Harvest Model.

Fisheries north of the border between Washington and Canada are modeled based on the framework under which they are managed.  Fisheries that are subject to Aggregate Abundance Based Management (<span title = "AABM: Aggregate Abundance Based Management">AABM</span>) per the provisions of the Pacific Salmon Treaty (<span title = "PST: Pacific Salmon Treaty">PST</span>) Agreement are managed based on catch limits set annually by the Pacific Salmon Commission (<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pacific_Salmon_Commission_(PSC)" target = "_blank" title = "PSC: Pacific Salmon Commission">PSC</a>). For Southeast Alaska, the total all gear (troll, net, sport) annual catch limit is determined based on the catch per unit effort observed during the winter troll fishery. Catch limits for the Canadian <span title = "AABM: Aggregate Abundance Based Management">AABM</span> fisheries (Northern <span title = "B.C.: British Columbia">B.C.</span> and <span title = "WCVI: West Coast Vancouver Island">WCVI</span> troll and sport) are derived annually from preseason “abundance indices” produced by the annual calibration of the <span title = "PSC: Pacific Salmon Commission">PSC</span> Chinook Model. Catch limits for all <span title = "AABM: Aggregate Abundance Based Management">AABM</span> fisheries are allocated to gear and time period based on recent year averages and modeled as quotas. Catch limits for the Canadian <span title = "AABM: Aggregate Abundance Based Management">AABM</span> fisheries are not available until the <span title = "PSC: Pacific Salmon Commission">PSC</span> Chinook Model calibration is completed, usually by April 1. For model runs conducted prior to this, Canadian <span title = "AABM: Aggregate Abundance Based Management">AABM</span> fisheries are modeled using fishery scalers from the previous year’s preseason model run, which can result in increased uncertainty during the early stages of the annual assessment process. For the remaining Canadian fisheries that are managed to the Individual Stock Based Management (<span title = "ISBM: Individual Stock Based Management">ISBM</span>) provisions of the <span title = "Pacific Salmon Treaty">PST</span> Agreement, fishery inputs are received from <span title = "ADFG: Alaska Department of Fish and Game">ADFG</span> and <span title = "DFO: Department of Fisheries and Oceans">DFO</span> via their respective <span title = "PSC: Pacific Salmon Commission">PSC</span> Chinook Technical Committee (<span title = "CTC: Chinook Technical Committee">CTC</span>) members. Thus, Chinook <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> relies on fishery estimates derived from the <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Pacific_Salmon_Commission_(PSC)" target = "_blank" title = "PSC: Pacific Salmon Commission">PSC</a> Chinook model to support <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> model inputs for Alaskan and most Canadian fisheries.  The <span title = "PSC: Pacific Salmon Commission">PSC</span> Chinook model has better stock representation in these northern fisheries and consequently is assumed to provide a better representation of fishing effort changes relative to the <a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Base_Period" target = "_blank" title = "Base period">base period</a>.

<span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> fishery input methods for retention fisheries includes quotas (as a fixed catch), fishery effort scalers (e.g., Puget Sound marine sport) or terminal fishery harvest rates used during <span title = "TAMM: Terminal Area Management Module">TAMM</span> processing (e.g., Puget Sound terminal net). Chinook non-retention fishery mortalities are primarily modeled using estimates of sublegal and legal-size encounters.

# 2. Post-season Model Usage

Although Chinook <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> is primarily used for U.S. pre-season fishery impact assessment, it can also be used in a <a href = https://framverse.github.io/fram_doc/calcs_glossary.html#Post-season" target = "_blank" title = "Post-season">post-season</a> context.  These post-season model runs are used for several purposes: as a tool to evaluate the model's performance by comparing model estimates to independently derived estimates, to evaluate the performance of the fishery management system towards meeting conservation objectives for key stocks, to develop stock-specific management objectives, and to assess stock status.

Post-season <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> model runs assess what occurred in previous years based onco-manager agreed-to observed input values and the associated estimated abundances. These runs use the “backwards” <span title = "BkFRAM: Backwards FRAM" >BkFRAM</span> utility that finds <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span> starting abundances using inputs of observed estimates for terminal run sizes and fishery catches (landed and non-retention). For Chinook <span title = "FRAM: Fishery Regulation Assessment Model">FRAM</span>, the postseason model runs are often referred to as "<a href = "https://framverse.github.io/fram_doc/calcs_glossary.html#Validation_Runs" target = "_blank" title = "Validation runs">validation runs</a>," and are usually run by U.S. salmon co-managers for U.S. domestic management evaluation every two years (in even numbered years; e.g., in 2020 for all fishery years 2018 and prior). Post-season model runs are often lagged by several years due to data availability issues.

<br>

**To cite this page:** <br>
`r paste("Salmon modeling and analysis workgroup.", format(Sys.Date(), "%Y."), page_title, "*in* FRAM Documentation. https://framverse.github.io/fram_doc/ ", paste("built", format(Sys.Date(), "%B %d, %Y.")))`
