```{r page title, include=FALSE}
page_title <- "Base Period Documentation - Appendix"
```

--- 
title: `r page_title`
author: ""
date: ""
description: "Getting to know the Fishery Regulation Assessment Model (FRAM)"
editor_options: 
  chunk_output_type: console
---




# 1. Appendix

## 1.1 Stock Profiles Jon

## 1.2 Fishery Profiles Oliver

## 1.3 Database Tables
All the input and output data for a calibration is located in an Access database titled "CalibrationSupport_xxxx", where the "xxxx" signifies a name/date that uniquely identifies the calibration. An effort was made to capture nearly all input data and manipulations in the database and program code for documentation and reproducibility purposes.

### Table Description
Input tables are 'blue', output tables are 'green', and legacy tables that are no longer used are 'white'.
```{r DBTables, echo = FALSE, fig.cap=c("Calibration Database Table Descriptions"),fig.pos='H', out.height = "800px", out.width='800px'}

knitr::include_graphics("literature/CalibrationDB Table Description.pdf")

```

#### [AEQs](#AEQ) 
 
```{r AEQTable,  results='asis'}
AEQTable<-as.data.frame(read.csv("objects/AEQ.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:AEQTable)",
                      "Adult Equivalent Values",
                      "</caption>"),
                      "</p>")
AEQTable

```

#### Base Period Cohort Sizes 

```{r BPCohortTable,  results='asis'}
BPCohortTable<-as.data.frame(read.csv("objects/BaseCohort.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:BPCohortTable)",
                      "Base Period Cohort Sizes",
                      "</caption>"),
                      "</p>")
BPCohortTable
```

#### Base Period Exploitation Rates

```{r BPRate,  results='asis'}
BPRate<-as.data.frame(read.csv("objects/BaseExploitationRate.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:BPRate)",
                      "Base Period Exploitation Rates",
                      "</caption>"),
                      "</p>")
BPRate

```

#### Base Period Marked Catches
```{r MrkCatch,  results='asis'}
MrkCatch<-as.data.frame(read.csv("objects/BasePeriodCatch.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:MrkCatch)",
                      "Base Period Marked Catches",
                      "</caption>"),
                      "</p>")
MrkCatch

```

#### Base Period Total Catches
```{r TotalCatch,  results='asis'}
TotalCatch<-as.data.frame(read.csv("objects/BasePeriodCatch_Tot.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:TotalCatch)",
                      "Base Period Total Catches (Marked + Unmarked)",
                      "</caption>"),
                      "</p>")
TotalCatch

```

#### Base Period Marked Escapements
```{r MrkEscape,  results='asis'}
MrkEscape<-as.data.frame(read.csv("objects/BPEscapements.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:MrkEscape)",
                      "Base Period Marked Escapements",
                      "</caption>"),
                      "</p>")
MrkEscape

```

#### Base Period Total Escapements
```{r TotalEscape,  results='asis'}
TotalEscape<-as.data.frame(read.csv("objects/BPEscapements_Tot.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:TotalEscape)",
                      "Base Period Total Escapements (Marked + Unmarked)",
                      "</caption>"),
                      "</p>")
TotalEscape

```

#### Coded-Wire-Tag Adjustments for Sinclair Inlet (Area 10E)
```{r Sinclair,  results='asis'}
Sinclair<-as.data.frame(read.csv("objects/CWT 10E adjustments.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:Sinclair)",
                      "Sinclair Inlet CWT Adjustments",
                      "</caption>"),
                      "</p>")
Sinclair
```

#### Out-of-Base CWTs for White River 
```{r WhiteOOB,  results='asis'}
WhiteOOB<-as.data.frame(read.csv("objects/CWT_WhiteOOB.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:WhieOOB)",
                      "White River Out-of-Base CWT Recoveries",
                      "</caption>"),
                      "</p>")
WhiteOOB
```

#### Coded-Wire-Tag Recovery Data
```{r CWTAll,  results='asis'}
CWTAll<-as.data.frame(read.csv("objects/CWTAll.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:CWTAll)",
                      "Coded-Wire-Tag Recovery Data",
                      "</caption>"),
                      "</p>")
CWTAll
```

#### Coded-Wire-Tags for South Puget Sound Surrogate Method
```{r CWTSPS,  results='asis'}
CWTSPS<-as.data.frame(read.csv("objects/CWTforSPS_SurroagteMethod.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:CWTSPS)",
                      "South Sound Coded-Wire-Tags",
                      "</caption>"),
                      "</p>")
CWTSPS
```

#### Coded-Wire-Tags to Impute Age 5 Recoveries
```{r CWTAge5,  results='asis'}
CWTAge5<-as.data.frame(read.csv("objects/CWTImputedAge5.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:CWTAge5)",
                      "Coded-Wire-Tags to Impute Age 5 Recoveries",
                      "</caption>"),
                      "</p>")
CWTAge5
```


#### Coded-Wire-Tags Recoveries for Strait of Juan de Fuca (JDF) 
```{r CWTJDF,  results='asis'}
CWTJDF<-as.data.frame(read.csv("objects/CWTJDFSurrogate071816.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:CWTJDF)",
                      "JDF Surrogate CWTs",
                      "</caption>"),
                      "</p>")
CWTJDF
```

#### Base Period Coded-Wire-Tags Recoveries Merged with Out-Of-Base Recoveries
```{r CWTMerged,  results='asis'}
CWTMerged<-as.data.frame(read.csv("objects/CWTOOBMerged.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:CWTMerged)",
                      "Base Period CWTs Merged with Out-Of-Base Recoveries",
                      "</caption>"),
                      "</p>")
CWTMerged
```

#### FRAM Chinook Fisheries
```{r Fish,  results='asis'}
Fish<-as.data.frame(read.csv("objects/Fishery.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:Fish)",
                      "JDF Surrogate CWTs",
                      "</caption>"),
                      "</p>")
Fish
```

#### Marked Fishery Model Stock Proportions 
```{r MSP,  results='asis'}
MSP<-as.data.frame(read.csv("objects/FisheryModelStockProportion.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:MSP)",
                      "Marked Model Stock Proportions",
                      "</caption>"),
                      "</p>")
MSP
```

#### Total (Marked + Unmarked) Fishery Model Stock Proportions 
```{r MSPTot,  results='asis'}
MSPTot<-as.data.frame(read.csv("objects/FisheryModelStockProportion_Tot.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:MSPTot)",
                      "Total (Marked + Unmarked) Model Stock Proportions",
                      "</caption>"),
                      "</p>")
MSPTot
```

#### Fishery Scalers
```{r FishScaler,  results='asis'}
FishScaler<-as.data.frame(read.csv("objects/FishScalar.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:FishScaler)",
                      "Fishery Scalers",
                      "</caption>"),
                      "</p>")
FishScaler
```

#### Von Bertalaffy Growth Parameters
```{r VBGrowth,  results='asis'}
VBGrowth<-as.data.frame(read.csv("objects/Growth.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:VBGrowth)",
                      "Von Bertalaffy Growth Parameters",
                      "</caption>"),
                      "</p>")
VBGrowth
```

#### Instructions for Imputing Surrogate CWT Recoveries
```{r CWTSurrogate,  results='asis'}
CWTSurrogate<-as.data.frame(read.csv("objects/ImputeRecoveries.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:CWTSurrogate)",
                      "Instructions for Imputing Surrogate CWT Recoveries",
                      "</caption>"),
                      "</p>")
CWTSurrogate
```

#### [Incidental Mortality Rates]{#Inc}
```{r Incidental,  results='asis'}
Incidental<-as.data.frame(read.csv("objects/IncidentalRate.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:Incidental)",
                      "Incidental Mortality Rates",
                      "</caption>"),
                      "</p>")
Incidental
```

#### Maturation Rates
```{r MatRate,  results='asis'}
MatRate<-as.data.frame(read.csv("objects/MaturationRate.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:MatRate)",
                      "Maturation Rates",
                      "</caption>"),
                      "</p>")
MatRate
```

#### Natural Mortality Rates
```{r NatRate,  results='asis'}
NatRate<-as.data.frame(read.csv("objects/NaturalMortality.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:NatRate)",
                      "Natural Mortality Rates",
                      "</caption>"),
                      "</p>")
NatRate
```

#### Marked Non-Retention Chinook Encounters
```{r NR,  results='asis'}
NR<-as.data.frame(read.csv("objects/NonRetention.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:NR)",
                      "Marked Non-Retention Chinook Encounters",
                      "</caption>"),
                      "</p>")
NR
```

#### Total (Marked + Unmarked) Non-Retention Chinook Encounters
```{r NRTot,  results='asis'}
NRTot<-as.data.frame(read.csv("objects/NonRetention_Tot.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:NRTot)",
                      "Total (Marked + Unmarked) Non-Retention Chinook Encounters",
                      "</caption>"),
                      "</p>")
NRTot
```

#### [Sublegal Mortality Rates]{#SubRate}
```{r SLRate,  results='asis'}
SLRate<-as.data.frame(read.csv("objects/ShakerMortRate.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:SLRate)",
                      "Sublegal Mortality Rates",
                      "</caption>"),
                      "</p>")
SLRate
```

#### Minimum Size Limits
```{r SizeL,  results='asis'}
SizeL<-as.data.frame(read.csv("objects/SizeLimits.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:SizeL)",
                      "Minimum Size Limits",
                      "</caption>"),
                      "</p>")
SizeL
```

#### FRAM Chinook Stocks
```{r Stocks,  results='asis'}
Stocks<-as.data.frame(read.csv("objects/Stock.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:Stocks)",
                      "Maturation Rates",
                      "</caption>"),
                      "</p>")
Stocks
```

#### Run Summary Information
```{r Run,  results='asis'}
Run<-as.data.frame(read.csv("objects/SummaryInfo.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:Run)",
                      "Run Summary Information",
                      "</caption>"),
                      "</p>")
Run
```

#### Surrogate Fishery Base Period Exploitation Rates
```{r SurrogateBPER,  results='asis'}
SurrogateBPER<-as.data.frame(read.csv("objects/SurrogateFishBPER.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:SurrogateBPER)",
                      "Surrogate Fishery Base Period Exploitation Rates",
                      "</caption>"),
                      "</p>")
SurrogateBPER
```

#### Target Encounter Rates
```{r TargetER,  results='asis'}
TargetER<-as.data.frame(read.csv("objects/TargetEncounterRates.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:TargetER)",
                      "Target Encounter Rates",
                      "</caption>"),
                      "</p>")
TargetER
```

#### Terminal Fishery Flags
```{r TermFlags,  results='asis'}
TermFlags<-as.data.frame(read.csv("objects/TerminalFisheryFlag.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:TermFlags)",
                      "Terminal Fishery Flags",
                      "</caption>"),
                      "</p>")
TermFlags
```

#### FRAM Time Steps
```{r TimeSteps,  results='asis'}
TimeSteps<-as.data.frame(read.csv("objects/TimeStep.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:TimeSteps)",
                      "FRAM Time Steps",
                      "</caption>"),
                      "</p>")
TimeSteps
```

#### Base Period Update Log
```{r Log,  results='asis'}
Log<-as.data.frame(read.csv("objects/UpdateLog.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:Log)",
                      "Base Period Update Log",
                      "</caption>"),
                      "</p>")
Log
```

#### CWT Weighting
```{r Weighting,  results='asis'}
Weighting<-as.data.frame(read.csv("objects/Weighting.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:Weighting)",
                      "CWT Weighting",
                      "</caption>"),
                      "</p>")
Weighting
```

#### White River Maturation Rates
```{r WhiteMatR,  results='asis'}
WhiteMatR<-as.data.frame(read.csv("objects/WhiteMatRate.csv"), header = TRUE)|>
 	 DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  rownames = FALSE,
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:WhiteMatR)",
                      "White River Maturation Rates",
                      "</caption>"),
                      "</p>")
WhiteMatR
```


## 1.4 Length-At-Age Distribution Parmeters

```{r LAge, results='asis'}
lengthtable<-as.data.frame(read.csv("objects/Length at Age.csv", header = TRUE))|>
  DT::datatable(filter = 'top',
                extensions = 'Buttons',
                options = list(
                  dom = 'Bfrtip',
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

cat("<p>", paste0("<caption>",
                      "(#tab:LAge)",
                      "Length-At-Age Distribution Parameters
Stage 1 analysis results. Maximum likelihood estimates of the mean, SD, and CV of fork length (FL, mm) by age (in months) for stock-time estimation strata for which sufficient records occurred to estimate distributional parameters (N = 20). Note, the two suspected outliers that were included in VBGF estimation but not in the CV function analysis are denoted by ‘**’; columns ‘Age’ and ‘Mean FL (cm)’ are the values used for fitting VBGFs.",
                      "</caption>"),
                      "</p>")

lengthtable
```


## 1.4 WinBUGS Code for Fitting Growth Functions {#WB}

```{r eval = FALSE, echo=TRUE}
##WinBUGS code for VB Growth fxn  #####
### Specs for FRAM VBGF parameter estimation: 
## N = 3 chains, thinned to 1 in 50
## N = 31k total interations (1k initial, 30k thereafter, summarize 5001+)
model 
{
	# priors
	mu.a0~dnorm(0, 0.001) # Mean hyperparameter for a0
	mu.Linf~dnorm(0, 0.001) # Mean hyperparameter for Linf
	mu.k~dnorm(0, 0.001) # Mean hyperparameter for k
	sigma.a0~dunif(0, 10000) # SD hyperparameter for a0
	sigma.Linf~dunif(0, 10000) # SD hyperparameter for Linf
	sigma.k~dunif(0, 10000) # SD hyperparameter for k
	sigma~dunif(0, 10000) # Residual standard deviation
	tau.a0 <- 1/(sigma.a0*sigma.a0)
	tau.Linf <- 1/(sigma.Linf*sigma.Linf)
	tau.k <- 1/(sigma.k*sigma.k)
	tau <- 1/(sigma*sigma) # Residual precision

	# hierarchical parmeters
	for (i in 1:P) {
   		Linf[i]~dnorm(mu.Linf, tau.Linf) 
   		k[i]~dnorm(mu.k, tau.k)#I(0,)
   		a0[i]~dnorm(mu.a0, tau.a0) 
	}

	# likelihood
	for (i in 1:N) {
		li[i] ~ dnorm(Lai[i],tau)
		Lai[i]<-Linf[pop[i]]*(1-exp(-k[pop[i]]*(ai[i]-a0[pop[i]])))  
		# Observation-level GOF calcs
		pred[i] <- Lai[i]
		rep_li[i] ~ dnorm(pred[i], tau)   #simulate perfect dataset
		rep_resid[i] <- rep_li[i]-pred[i]   #calc resid for simulated data
		resid[i] <- li[i] - pred[i]   #calc resid for actual data
		sq[i] <- pow(resid[i],2)  #calc squared resids for actual data 								sq_new[i] <- pow(rep_resid[i],2)   #calc squared resids for simulated data
	}
	# Dataset-level GOF calcs
	fit<-sum(sq[]) #sum squared resids for actual data
	fit_new<-sum(sq_new[]) #sum squared resids for simulated data
	test<-step(fit_new-fit) #determine which is greater (0s and 1s)
	bpvalue<-mean(test) #mean of 0,1 gives Bayesian P-value
}	

# inits
list(a0=c(-5,-5,-5,-5,-5,-5,-5,-5,-5,-5,-5,-5,-5), k=c(0.03,0.03,0.03,0.03,0.03,0.03,0.03,0.03,0.03,0.03,0.03,0.03,0.03), Linf=c(950,950,950,950,950,950,950,950,950,950,950,950,950),mu.a0=-5,mu.k=0.03,mu.Linf=950,sigma.a0=1,sigma.Linf=1,sigma.k=1,sigma=1)

#data -- also must load secondary data file (Appendix A) with stock-month means
list(N=164, P=13) #n is 165 observations (0 indexed) and 13 stock aggregates


```